---
title: "Peak integration script"
author: "Jakob Koch"
date: "18/02/2020"
output: html_notebook
---

**history of the versions...**
Fri Nov 22 14:12:30 2019
Tue Feb 18 11:02:09 2020 
Tue Mar 24 12:48:23 2020 
Mon Apr 06 11:28:46 2020 last

# Setup
```{r setup}
# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
if (!require("pacman")) install.packages("pacman")
pacman::p_load(mzR,here,tidyverse,furrr,writexl,readxl)
here()
# library(minpack.lm)
# library(tidyverse)
# setwd("~/Version_control_test/phospholipid_ms2_classifier/Project_fromHD/MAK_Peakextractor")
source(file = here("Functions","Functions_1_peak_integration_automated_Dilutions.R"))
# readxl,pheatmap,writexl
```

## Load data
the mzML files are defined as by their ending and an additional user defined filter.
one should here select only one kind of file, either *Sample* or *Dilution* alike types


*Bitte filtere von der Excelpeaklist nur entweder 
die Dilutions (defined peaklist) raus (für diese werden die ersten beiden eXcel Tabellenblätter verwendet,
oder alle Samples (large peaklist), raus und rechne das entsprechend....
ich sollte da längst eine Anleitung geschrieben haben....*

* different peaklists can be combined
* different algorithms shouldn't be combined/used for one batch of data.


```{r Load Data1}
      # directory <- "O:/LCMS/PLAN/2019/20190808_PLAN032_HAP1pos_Naschberger_Ruediger_GPX4_oxPL/Alexander_HEK_HSD10/HEK_HSD10_CL/Raw Data"
      # directory <- "Z:/LCMS/PLAN/2019/20190930PLAN033_AHuber_EWerner_GPX4/E_Werner_mouse_tissues"
      # directory new folder layout usage automated mzML data finding enabled:

# Readout raw data file information ---------------------------------------

Project_mzML.files <- readxl::read_xlsx(path = here(paste0("Proj_mzMLs.xlsx")))
View(Project_mzML.files)
Project_mzML.files <- Project_mzML.files %>% filter(str_detect(string = Filename,pattern = "DIV"))
Project_mzML.files <- Project_mzML.files %>% filter(str_detect(string = Filename,pattern = ".*ution"))
# 
# Project_mzML.files <- Project_mzML.files %>% filter(!str_detect(string = Filename,pattern = ".*ution"))
# Project_mzML.files <- Project_mzML.files %>% filter(!str_detect(string = Filename,pattern = ".*QC"))

filelist <- Project_mzML.files$Filepath
filelist

filelist2 <- (str_extract(string = filelist,pattern = "[^/]+$") %>% str_split(pattern = "\\.",simplify = T))[,1]
filelist2
```

## Important!
The pattern here has to match either 

* Samples or
* Dilutions
for script to work propperly!

```{r Selection of Folders,eval=FALSE}
folder <- "1_Dilutions_peakf2_"
folder <- "2_Samples_peakf3_"
folder <- "2_QCs"

```

## Load Peaklists
In a second step the rtstdlist, a list containing Std peaks for RT correction is loaded. (Manual input is still present, but unused)
Here the peaklists are loaded in a list format, one list entry per XLSX page.
Comments are saved to an additional vector *comments* for later use in the single lipid plots.
```{r Load Data2}
peaklist_file <- choose.files(default = here("2_PL_peaklist.xlsx"),multi = F,index = 1,filters = matrix(c("xlsx","*.xlsx","csv","*.csv"),ncol=2,byrow = T))
peaklists <- list()
for (i in seq_along(excel_sheets(peaklist_file))) {
  peaklists[[i]] <- read_xlsx(path = peaklist_file,sheet = i) 
}
names(peaklists) <- excel_sheets(peaklist_file)

names(peaklists)


if(str_detect(string = folder,pattern = "Samples")) {
  # samples:
  lipidlist <- peaklists$PLOP1 %>% select(1:3)
  comments <- peaklists$PLOP1 %>% select(lipid,4:ncol(.))
  rtstdlist <- peaklists$PLOP2 %>% select(1:3)
} else {
  # dilutions
  lipidlist <- peaklists$Dilutions %>% select(1:3)
  rtstdlist <- peaklists$Rtstdlist_Dilutions %>% select(1:3)
  comments <- peaklists$Dilutions %>% select(lipid,4:ncol(.))
}

# check for duplicate lipid id
lipidlist %>% group_by(lipid) %>% summarise(N=n()) %>% filter(N>1)
```

# LOOP for automated Readout
Can easily be run as a independent R background job, works great, just mark from line 132 onwards.
```{r -# Initialize result object}
dir.create(path = here(paste0(folder,Sys.Date())))
Folder_created <- paste0(folder,Sys.Date())
# Folder_created <- paste0(folder,"2019-12-11")
progress_bar <- progress_estimated(n = length(filelist))
 # i=1

# For loop, does everything updated-------------------------------------------
# Fri Sep 06 14:00:46 2019 ------------------------------

pacman::p_load(mzR,here,tidyverse,furrr,writexl,readxl)
source(file = here("Functions","Functions_1_peak_integration_automated_Dilutions.R"))
# for(i in 12){
rtstdlist1 <- rtstdlist
  i=1  
  print(paste("estimated runtime ~", length(filelist)*5, "min"))
for(i in seq_along(filelist)){
  # Dilutions_Integration <- function(filelist,filelist2,rtstdlist,lipidlist) {}
  print(paste(i, "of", length(filelist)))
  file_name <- filelist2[i]
  print(file_name)
  mz <- openMSfile(filelist[i])
  
  ms1_header <- header(mz) %>% filter(msLevel==1)
  ms2_header <- header(mz) %>% filter(msLevel==2)
  rtstdlist1 -> rtstdlist
  # extracts calibration data
  cal1 <- apply(X = rtstdlist1, MARGIN = 1, 
                FUN = function(x) calibrate(x=x, 
                                            mz=mz, 
                                            ms1_header=ms1_header))
  # filter for max peak time!
  # here one needs to summarize, in the rare event of two MS1 signals at the peak maximum...
  extracted_cal_info <- bind_rows(
    map(cal1,.f = function(x){
      x %>% filter(Peak_max==T) %>%summarise(rt_times=mean(rt_times),m_z=mean(m_z),Intensity=mean(Intensity)) %>% select(rt_times,m_z,Intensity)
    })
  )
  rtstdlist <- bind_cols(rtstdlist1,extracted_cal_info) %>% rename(found="rt_times",mz_found="m_z") %>% filter(Intensity >=max(Intensity)*0.1)# %>% select(-Intensity)
  
  cal_model <- lm(y~x,data = data.frame(x=rtstdlist$rt, y=rtstdlist$found),weights = rtstdlist$Intensity)
  rtstdlist$rt_corr  <- predict(cal_model, newdata=data.frame(x=rtstdlist$rt))
  lipidlist$rt_corr <- predict(cal_model, newdata=data.frame(x=lipidlist$rt))
  rtstdlist$Sample <- file_name

  # use this line to export the found and corrected cal. information -> future save rtstdlist & lipidlist into Rdata
  # writexl::write_xlsx(x = rtstdlist,path = here(paste("New_automatically_extracted_Std_list_",Sys.Date(),".xlsx")))
  # write.csv(x = rtstdlist %>% select(mz_found,found,lipid),file = here(paste("New_automatically_extracted_Std_list_",Sys.Date(),".csv")),row.names = F)
  p_cal <- ggplot(data =  bind_rows(cal1),mapping = aes(x = rt_times,y = Intensity))+
    geom_line()+
    scale_alpha_continuous(range = c(0,1))+
    geom_vline(data = rtstdlist,mapping = aes(xintercept = rt,colour="Searched_at"),linetype=5)+
    geom_vline(data = rtstdlist,mapping = aes(xintercept = found,colour="Found_at"),linetype=5)+
    geom_vline(data = rtstdlist,mapping = aes(xintercept = rt_corr,colour="Corrected"),linetype=5)+
    scale_color_manual(name="Lines",values = c("Searched_at"="red","Found_at"="blue","Corrected"="green"))+
    facet_wrap(~lipid,scales = "free")+
    labs(title = file_name)+
    theme_light()+
    theme(plot.title = element_text(hjust = 0.5))
  ggsave(filename = here(Folder_created,paste0(file_name,"_cal_",Sys.Date(),".pdf")),plot = p_cal,device = "pdf",width = 11.7,height = 8.3)
  
  
  # old apply approach working...
   tictoc::tic()
   peak_res <- apply(X = lipidlist, MARGIN = 1,
                     FUN = function(x) peakextrakt2(rt=as.numeric(x[4]),
                                                   mass=as.numeric(x[3]),
                                                   mz=mz,
                                                   ms1_header=ms1_header
                                                   ,ms2_header=ms2_header
                                                   ))
   tictoc::toc()
   
   
   Peaklist <- bind_cols(lipidlist %>% rename_at(.vars = c(2:4),.funs = function(x) paste0("A_",x)),
                         bind_rows(peak_res))  %>% mutate(Sample=filelist2[i]) %>% select(Sample,everything())
   tictoc::toc()

  
   window <- c(500,1000) #lower and upper cut off
  
   rt_min <-  ms1_header[which.min(abs(ms1_header$retentionTime - 0*60)), 1]
   rt_max <-  ms1_header[which.min(abs(ms1_header$retentionTime - 12*60+12)), 1]

# CL_Version 0.1 JK   
#  window <- c(1000,1650) #lower and upper cut off 
#  rt_min <-  ms1_header[which.min(abs(ms1_header$retentionTime - 6*60)), 1]
#  rt_max <-  ms1_header[which.min(abs(ms1_header$retentionTime - 22*60)), 1]
  
   rt_spectra <- ms1_header$seqNum[ms1_header$seqNum >= rt_min & ms1_header$seqNum <= rt_max]
   rt_times <- ms1_header$retentionTime[ms1_header$seqNum >= rt_min & ms1_header$seqNum <= rt_max]
  
   # Data extraction for the 2D plots
   spectra <- peaks(mz, rt_spectra)
   Spectra <- future_map(spectra,.f = function(x) {tibble(m_z=x[,1],Intensity=x[,2])})
   Spectra <- bind_rows(bind_rows(setNames(Spectra, seq_along(Spectra)), .id = "id")) %>% rename(Spectrum= id)
  
   # object <- tibble(Spectra=spectra) %>% mutate(Spectra=future_map(Spectra,.f = function(x) {tibble(m_z=x[,1],Intensity=x[,2])}))
   # head(object)
   # chrom_plot <- object %>% unnest(.id = "Spectrum")
  
   chrom_plot <- left_join(x = Spectra,
                           y = tibble(rt_times2=rt_times/60) %>% rownames_to_column(var = "Spectrum"),
                           by="Spectrum")
  
  
   # HIER PASSIERT DAS FILTERING1
   rect_data <- (Peaklist %>% select(lipid,starts_with(match = "C")))
   rect_data2 <- (Peaklist %>% select(lipid,starts_with(match = "D")))
   rect_data_p <- rect_data %>% filter(C_mz_low > window[1] & C_mz_high < window[2])
  
  
  
     plotting1 <- chrom_plot %>% filter(m_z < window[2]+1 & m_z > window[1]-1)
   # plotting$Intensity
   cuter <- 100
   plotting <- plotting1 %>% filter(Intensity > cuter)
   p2D <- ggplot()+
     geom_hex(data = plotting, mapping = aes(x = rt_times2,y = m_z,weight=log2(Intensity)),stat = "binhex",binwidth=c(.1,.2))+ # binwidth(x,y)
     # scale_fill_gradient2(low = "white",mid = "blue",high = "black",midpoint = max(log10(plotting$Intensity))*10,na.value = "white")+
     # scale_fill_gradient(low = "white",high = "red",na.value = "white")+
     scale_fill_gradient(low = "white",high = "black",na.value = "white")+
     geom_rect(data = rect_data_p,mapping = aes(xmin=C_rt_Start,xmax=C_rt_End,ymin=C_mz_low, ymax=C_mz_high),fill="red",alpha=0.8)+
     geom_point(data = lipidlist,mapping = aes(x=rt_corr,y = mz),size=0.1,alpha=0.9,col="blue")+
     geom_point(data = rect_data2,mapping = aes(x=D_rt_times,y = D_m_z),size=.1,alpha=0.8,col="green")+
     geom_text(data = rect_data2,mapping = aes(x=D_rt_times,y = D_m_z+1,label=lipid),size=0.2,alpha=0.8,col="black")+
     # scale_fill_grey()na.value = "white",low = "yellow",mid = "blue",high = "black",midpoint = 3)+
     scale_x_continuous(limits = c(0,12),breaks = function(limits) {seq(from=limits[1],to = limits[2],by = 1)},minor_breaks = function(limits) {seq(from=limits[1],to = limits[2],by = 0.2)},expand = c(0,0))+
    scale_y_continuous(limits = c(500,950),breaks = function(limits) {seq(from=limits[1],to = limits[2],by = 5)},minor_breaks = function(limits) {seq(from=limits[1],to = limits[2],by = 1)},expand = c(0,0))+
    theme_bw()+
    theme(legend.position = "none",axis.text = element_text(size = 6))
  ggsave(filename = here(Folder_created,paste0(filelist2[i],"_2D_",Sys.Date(),".pdf")),plot = p2D,device = "pdf",width = 16.5,height = 11.7)

  close(mz)
 
  # version for Samples:
  # reintroduce comments:
    Peaklist2 <- left_join(Peaklist,y = comments,by="lipid")
    ypsilon <- Peaklist2 %>% arrange(lipid) %>% transmute(lipid1=lipid,lipid,comment,info) %>% group_by(lipid1) %>%  nest()

    ## part to create single plots per peak and save to named pdf per sample:
    xilon <- (Peaklist %>% arrange(lipid))$Peakshape
  ###create the single plots - works only for samples with comments,
  temp <- map2(.x = xilon,.y = ypsilon$data,.f = function(x,y){
    ggplot(data = x[[1]],mapping = aes(x = rt_times,y = Intensity))+
      geom_line()+
      geom_area(data = x[[2]],mapping = aes(x = rt_times,y = Intensity,fill="red",alpha=.6))+
      labs(title = y[[1]],caption = paste(y[[2]],y[[3]],sep = "; "))+theme_bw()+theme(legend.position = "none")
  })
  ggpubr::ggexport(gridExtra::marrangeGrob(grobs = temp,nrow=2,ncol=2),filename = here(Folder_created,paste0(filelist2[i],"_single_plots_",Sys.Date(),".pdf")))
    # 
    # # version for dilutions.
    #       # reintroduce comments:
    #       Peaklist <- left_join(Peaklist,y = comments,by="lipid")
    #       ypsilon <- Peaklist %>% arrange(lipid) %>% transmute(lipid1=lipid,lipid,comment,info) %>% group_by(lipid1) %>%  nest()
    #       
    #       ## part to create single plots per peak and save to named pdf per sample:
    #       xilon <- (Peaklist %>% arrange(lipid))$Peakshape
    #     
    #     # create the single plots
    #     temp <- map2(.x = xilon,.y = ypsilon$data,.f = function(x,y){
    #   
    #       ggplot(data = x[[1]],mapping = aes(x = rt_times,y = Intensity))+
    #         geom_line()+
    #         geom_area(data = x[[2]],mapping = aes(x = rt_times,y = Intensity,fill="red",alpha=.6))+
    #         labs(title = y[[1]])+theme_bw()+theme(legend.position = "none")
    #     })
    #     
    #     # temp <- Peaklist %>% arrange(lipid) %>% mutate(single_Plots=map2(.x = Peakshape,.y = ypsilon,.f = function(x,y){
    #     #   ggplot(data = x[[1]],mapping = aes(x = rt_times,y = Intensity))+
    #     #     geom_line()+
    #     #     geom_area(data = x[[2]],mapping = aes(x = rt_times,y = Intensity,fill="red",alpha=.6))+
    #     #     labs(title = y[[1]],caption = str_c(y[[2]],y[[3]]))+theme(legend.position = "none")
    #     # }))
    #     ggpubr::ggexport(gridExtra::marrangeGrob(grobs = temp,nrow=2,ncol=2),filename = here(Folder_created,paste0(filelist2[i],"_single_plots_",Sys.Date(),".pdf")))
      
        save(Peaklist,file = here(Folder_created,paste0(filelist2[i],"_1_",Sys.Date(),".Rdata")))
          gc()
          progress_bar$tick()$print()
}
```